Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-25 cav2017/cav2017.tex
minor change

We describe and evaluate two approaches for synthesizing digital controllers for state-space physical plants, 
both based on the Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} architecture, recently of wide use for program synthesis.   
We provide two alternative automated procedures to synthesize safe 
digital state-feedback controls
and prove their soundness by encompassing errors due to digitalization and quantization effects that are present in the digital controller interacting with the continuous plant. 
% computations and representations of dynamics, providing sufficient conditions for sound and complete safe solutions for the state feedback control problem.

@@

Both models synthesize sound safe controllers given a discrete-time model of the closed-loop system.

We provide experimental results 
showing that both approaches are able to efficiently synthesize stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is $7.9$ seconds, and most controllers can be synthesized in less
than $17.2$ seconds. We further show that, in a direct comparison, the abstraction-based
approach lowers the median runtime of our benchmarks by a factor of seven over the
multi-staged approach.

@@

\begin{enumerate}
%
\item In this paper we extend the problem of digital control synthesis 
  from a transfer function representation into a state space representation. 
  Such representation allows to study the  input-to-state dynamics, while 
  transfer functions represents only input-output relations and cannot be 
  enough to ensure deeper properties, e.g., the safety property.
%
\item We proposes a synthesis of state feedback controllers that ensures 
  safety properties, that is only  possible in state-space representation. 
  In addition to the already existing stability conditions which is a much 
  harder problem.
%
\item We develop two approaches: the first, naive, one relies on
  explicit unfolding of the system's dynamics up to the completeness
  threshold, whilst the second one is abstraction-based and leverages
  the concepts of abstraction refinement and acceleration to increase 
  speed whilst retaining soundness. Both approaches provide sound
  synthesis of state feedback systems considering the various
  error sources in the implementation of the control algorithm and in the
  computational modelling of plant dynamics.
%
\item We develop a model for different sources quantization and their
  effect on the state space progression, for which we find safe bounds
  that ensure the safety of our controllers in a hybrid
  continuous-digital domain.
%
\end{enumerate}

@@

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding errors
during computation.  In~\cite{fialho1994stability}, a stability measure
based on the error of the digital dynamics ensures that the deviation
introduced by FWL does not make the digital system unstable.  A~more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  The analyses in~\cite{DBLP:conf/hybrid/RouxJG15,
DBLP:conf/hybrid/WangGRJF16} rely on an invariant computation on the
discrete system dynamics using Semi-Definite Programming (SDP).  While the
former uses bounded-input and bounded-output (BIBO) properties to determine stability, 
the latter uses Lyapunov-based quadratic invariants.  In both cases, the SDP solver uses
floating-point arithmetic and soundness is checked by bounding the error. 
An alternative approach is taken by~\cite{park2016scalable}, where the
verification of given control code is performed against a known model by
extracting an LTI model of the code through symbolic execution.  In order to
account for rounding errors, an upper bound is introduced in the
verification phase.  The work in \cite{picasso2003stabilization,
picasso2002construction} introduces attractive invariant sets as a mechanism
to bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set.  In a similar fashion,
\cite{liberzon2003hybrid} evaluates the dynamics of the quantization error
and bounds its trajectory to a known region over a finite period of time. 
This technique works for both linear and non-linear systems.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and 
zeros positions, degrading the closed-loop dynamics, causing steady-state 
errors (see the Appendix~\ref{sec:appendix:LTIbackground} for details) 
and eventually destabilizing  the system~\cite{Bessa16}. However, since 
in this paper we require stability only as a precursor to safety, it is 
sufficient to check that the (perturbed, noisy) model converges to a neighbourhood of the equilibrium within the safe set.

@@

At this point, we recap the the synthesis problem that we solve in this work:
we seek a digital feedback controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop plant model safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range, 
the reference signal to be set to zero, 
saturations on the inputs, 
and account for digitalization and quantization errors introduced by the controller.

@@

%\todo{please eventually itemize next}
\begin{enumerate}
\item The inductive synthesis phase (i.e., {\sc synthesize}) uses BMC to
compute a candidate solution $K$
that satisfies both the stability criteria detailed in 
Sec.~\ref{ssec:stability} and the safety specification in Sec.~\ref{ssec:safety}.
To synthesize a controller that satisfies the stability criteria, we require that
a computed characteristic polynomial satisifies Jury's criteria. The details of this
calculation can be found in the Appendix.

@@

\begin{algorithm}[]
\scriptsize
\begin{algorithmic}[1]
\Function{$safetyCheck()$}{}
\State assert($ \underline{u}  \leq u \leq \overline{u}$)
 \State set $x_0$ to be a vertex state, e.g., $[\underline{x_0},\underline{x_0}]$	
\For ($c=0; c < 2^{Num_States}; c++$)
  %\State assert($ \underline{x_0}  \leq x_0 \leq \overline{x_0}$)
	\While{$i=0; i< k; i++$}
		%\State $u = (\langle I_p,F_p\rangle)((\langle I_c,F_c\rangle)K * (\langle I_c,F_c\rangle) x)$
		\State $u = (plant\_typet)((controller\_typet)K * (controller\_typet) x)$
		\State $x = A * x + B * u$
		\State assert($\underline{x} \leq x \leq \overline{x}$ )
  	\EndWhile
  	\State set $x_0$ to be a new vertex state
  	\EndFor
  %\State Return $K$
\EndFunction
\end{algorithmic}
\label{alg:safetycheck}
\end{algorithm}

@@

\subsubsection{Illustrative Example}
%
Let us take the Helicopter example in Table \ref{table:State-space-benchmarks} with discretized dynamics
$$A_d = \left[\begin{array}{ccc}2.6207&-1.1793&0.65705\\2&0&0\\0&0.5&0\end{array}\right], B_d = \left [\begin{array}{c}8\\0\\0\end{array}\right]$$
Using the initial state bounds $\underline{x_{0}}=-0.9$ and $\overline{x_{0}}=0.9$, the input bounds $\underline{u}=-10$ and $\overline{u}=10$, and safety specifications 
$\underline{x_{i}}=-0.92$ and $\overline{x_{i}}=0.92$, the {\sc synthesize} phase in Fig.~\ref{fig:CEGARIS} synthesizes an initial candidate controller $K=[\begin{array}{ccc}0.24609375&-0.125&0.1484375\end{array}]$.
Although this candidate is stable, the {\sc verify} phase finds it to be unsafe and returns a list of iterations with an example initial state that fails the safety specification
$(k,x_0) \in \{ (2, [\begin{array}{ccc}0.9&-0.9&0.9\end{array}]), (3, [\begin{array}{ccc}0.9&-0.9&-0.9\end{array}])\}$.
This allows the {\sc abstract} phase to create a new safety specification that considers these iterations for these initial states to constrain the solution space.
This refinement allows {\sc synthesize} to find a controller $K=[\begin{array}{ccc}0.23828125&-0.17578125&0.109375\end{array}]$, which this time passes the verification phase, resulting in a safe system.

@@

We give the runtimes required to synthesize a stable controller for each
benchmark in Table~\ref{tab:results}.  
Here \textit{Benchmark} is the name of the respective
benchmark, $\mathcal{F}_{\langle I_p,F_p \rangle}$ is the fixed-point precision used
to model the plant, while \textit{Time}
is the total time (in milliseconds) required to synthesize a stable controller
for the given plant, for CEGIS with multi-staged verification as well as
abstraction-based CEGIS. The precision for the controller,
$\mathcal{F}_{\langle I_c,F_c \rangle}$, is always chosen to be $I_c = 8, F_c = 8$.

For the majority of our benchmarks, the conjecture stands that the abstraction-based back-end
is faster than the basic multi-staged verification approach. However, it does fail to
find solutions to 3 benchmarks, as compared to 2 with the multi-staged back-end, where
the numbers given in the discretization matrices are too small. In direct comparison, the
abstraction-based approach is on average able to find a solution in approximately 35\% of
the time required using the multi-staged back-end, and has a median runtime 1.15\,s, which is
seven times smaller than the multi-staged approach. The two back-ends complement each other
in benchmark coverage and together solve all benchmarks in the set.

@@

The median runtime for our benchmark set is 7.9\,s.  Overall,
the average synthesis time amounts to approximately 17.2\,s.  We
consider these times short enough to be of practical use to control
engineers, and thus affirm RQ1.

There are several cases for which the system
may fail to find a controller. For the naive approach, 
the completeness threshold may be too large, thus causing a
timeout.  On the other hand, the abstraction-based approach may
require a very precise abstraction, resulting in too many refinements
and, consequently, a timeout.  Yet another source of incompleness is the inability
of the {\sc synthesize} phase to use a large enough precision for the
plant model in a given problem.

@@

We have derived a number of benchmarks from the control literature, which we report below. 
As discussed, in this paper, we disregard the presence of output signals, that is the effect of matrices $\textbf{C}, \textbf{D}$. 
%

>>>>

We describe and evaluate two approaches for synthesizing digital controllers for state-space physical plants, 
both based on the Counter-Example Guided Inductive Synthesis (CEGIS)~\cite{sketch} architecture, recently of wide use for program synthesis.   
We provide two alternative automated procedures to synthesize safe 
digital state-feedback controls
and prove their soundness by encompassing errors due to digitalization and quantization 
effects that are present in the digital controller interacting with the continuous plant. 
Both models synthesize sound safe controllers given a discrete-time model of the closed-loop system.

@@

We provide experimental results 
showing that both approaches are able to efficiently synthesize stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median run-time for our benchmark
set is $7.9$ seconds, and most controllers can be synthesized in less
than $17.2$ seconds. We further show that, in a direct comparison, the abstraction-based
approach lowers the median run-time of our benchmarks by a factor of seven over the
multi-staged approach.

@@

\begin{enumerate}
%
\item In this paper we extend the problem of digital control synthesis 
  from a transfer function representation into a state space representation. 
  Such representation allows to study the  input-to-state dynamics, while 
  transfer functions represents only input-output relations and cannot be 
  enough to ensure deeper properties, e.g., the safety property.
%
\item We proposes a synthesis of state feedback controllers that ensures 
  safety properties, that is only  possible in state-space representation. 
  In addition to the already existing stability conditions which is a much 
  harder problem.
%
\item We develop two approaches: the first, naive, one relies on
  explicit unfolding of the system's dynamics up to the completeness
  threshold, whilst the second one is abstraction-based and leverages
  the concepts of abstraction refinement and acceleration to increase 
  speed whilst retaining soundness. Both approaches provide sound
  synthesis of state feedback systems considering the various
  error sources in the implementation of the control algorithm and in the
  computational modeling of plant dynamics.
%
\item We develop a model for different sources quantization and their
  effect on the state space progression, for which we find safe bounds
  that ensure the safety of our controllers in a hybrid
  continuous-digital domain.
%
\end{enumerate}

@@

There are two different problems that arise from FWL semantics.  The first
is the error in the dynamics caused by the inability to represent the exact
state of the physical system, while the second relates to rounding errors
during computation.  In~\cite{fialho1994stability}, a stability measure
based on the error of the digital dynamics ensures that the deviation
introduced by FWL does not make the digital system unstable.  A~more recent
approach~\cite{DBLP:journals/automatica/WuLCC09} uses $\mu$ calculus to
directly model the digital controller so that the selected parameters are
stable by design.  The analysis in~\cite{DBLP:conf/hybrid/RouxJG15,
DBLP:conf/hybrid/WangGRJF16} rely on an invariant computation on the
discrete system dynamics using Semi-Definite Programming (SDP).  While the
former uses bounded-input and bounded-output (BIBO) properties to determine stability, 
the latter uses Lyapunov-based quadratic invariants.  In both cases, the SDP solver uses
floating-point arithmetic and soundness is checked by bounding the error. 
An alternative approach is taken by~\cite{park2016scalable}, where the
verification of given control code is performed against a known model by
extracting an LTI model of the code through symbolic execution.  In order to
account for rounding errors, an upper bound is introduced in the
verification phase.  The work in \cite{picasso2003stabilization,
picasso2002construction} introduces attractive invariant sets as a mechanism
to bound the quantization error effect on stabilization as an invariant set
that always converges toward the controllable set.  In a similar fashion,
\cite{liberzon2003hybrid} evaluates the dynamics of the quantization error
and bounds its trajectory to a known region over a finite period of time. 
This technique works for both linear and non-linear systems.

@@

Previous studies~\cite{gangli1} show that the FWL affects the poles and 
zeros positions, degrading the closed-loop dynamics, causing steady-state 
errors (see the Appendix~\ref{sec:appendix:LTIbackground} for details) 
and eventually destabilizing  the system~\cite{Bessa16}. However, since 
in this paper we require stability only as a precursor to safety, it is 
sufficient to check that the (perturbed, noisy) model converges to a neighborhood of the equilibrium within the safe set.

@@

At this point, we recap the the synthesis problem that we solve in this work:
we seek a digital feedback controller $K$ (see Eq.~\ref{eq:controlaction})
that makes the closed-loop plant model safe for 
initial state $x_0$, reference signal $r_k$ and input $u_k$
as defined in Sec.~\ref{ssec:safety}. 
We consider non-deterministic initial states within a specified range, 
the reference signal to be set to zero, 
saturation on the inputs, 
and account for digitalization and quantization errors introduced by the controller.

@@

%\todo{please eventually itemize next}
\begin{enumerate}
\item The inductive synthesis phase (i.e., {\sc synthesize}) uses BMC to
compute a candidate solution $K$
that satisfies both the stability criteria detailed in 
Sec.~\ref{ssec:stability} and the safety specification in Sec.~\ref{ssec:safety}.
To synthesize a controller that satisfies the stability criteria, we require that
a computed characteristic polynomial satisfies Jury's criteria. The details of this
calculation can be found in the Appendix.

@@

\begin{algorithm}[]
\scriptsize
\begin{algorithmic}[1]
\Function{$safetyCheck()$}{}
\State assert($ \underline{u}  \leq u \leq \overline{u}$)
 \State set $x_0$ to be a vertex state, e.g., $[\underline{x_0},\underline{x_0}]$	
\For {($c=0; c < 2^{Num_States}; c++$)}
  %\State assert($ \underline{x_0}  \leq x_0 \leq \overline{x_0}$)
	\While{$i=0; i< k; i++$}
		%\State $u = (\langle I_p,F_p\rangle)((\langle I_c,F_c\rangle)K * (\langle I_c,F_c\rangle) x)$
		\State $u = (plant\_typet)((controller\_typet)K * (controller\_typet) x)$
		\State $x = A * x + B * u$
		\State assert($\underline{x} \leq x \leq \overline{x}$ )
  	\EndWhile
  	\State set $x_0$ to be a new vertex state
  	\EndFor
  %\State Return $K$
\EndFunction
\end{algorithmic}
\label{alg:safetycheck}
\end{algorithm}

@@

\subsubsection{Illustrative Example}
%
Let us take the Helicopter example in Table \ref{table:State-space-benchmarks} with discretized dynamics
$$A_d = \left[\begin{array}{ccc}2.6207&-1.1793&0.65705\\2&0&0\\0&0.5&0\end{array}\right], B_d = \left [\begin{array}{c}8\\0\\0\end{array}\right]$$
Using the initial state bounds $\underline{x_{0}}=-0.9$ and $\overline{x_{0}}=0.9$, the input bounds $\underline{u}=-10$ and $\overline{u}=10$, and safety specifications 
$\underline{x_{i}}=-0.92$ and $\overline{x_{i}}=0.92$, the {\sc synthesize} phase in Fig.~\ref{fig:CEGARIS} synthesizes an initial candidate controller $K=[\begin{array}{ccc}0.24609375&-0.125&0.1484375\end{array}]$.
Although this candidate is stable, the {\sc verify} phase finds it to be unsafe and returns a list of iterations with an example initial state that fails the safety specification
{\footnotesize $(k,x_0) {\in} \{ (2, [\begin{array}{ccc}0.9&-0.9&0.9\end{array}]), (3, [\begin{array}{ccc}0.9&-0.9&-0.9\end{array}])\}$}.
This allows the {\sc abstract} phase to create a new safety specification that considers these iterations for these initial states to constrain the solution space.
This refinement allows {\sc synthesize} to find a controller {\footnotesize $K=[\begin{array}{ccc}0.23828125&-0.17578125&0.109375\end{array}]$}, which this time passes the verification phase, resulting in a safe system.

@@

We give the run-times required to synthesize a stable controller for each
benchmark in Table~\ref{tab:results}.  
Here \textit{Benchmark} is the name of the respective
benchmark, $\mathcal{F}_{\langle I_p,F_p \rangle}$ is the fixed-point precision used
to model the plant, while \textit{Time}
is the total time (in milliseconds) required to synthesize a stable controller
for the given plant, for CEGIS with multi-staged verification as well as
abstraction-based CEGIS. The precision for the controller,
$\mathcal{F}_{\langle I_c,F_c \rangle}$, is always chosen to be $I_c = 8, F_c = 8$.

For the majority of our benchmarks, the conjecture stands that the abstraction-based back-end
is faster than the basic multi-staged verification approach. However, it does fail to
find solutions to 3 benchmarks, as compared to 2 with the multi-staged back-end, where
the numbers given in the discretization matrices are too small. In direct comparison, the
abstraction-based approach is on average able to find a solution in approximately 35\% of
the time required using the multi-staged back-end, and has a median run-time 1.15\,s, which is
seven times smaller than the multi-staged approach. The two back-ends complement each other
in benchmark coverage and together solve all benchmarks in the set.

@@

The median run-time for our benchmark set is 7.9\,s.  Overall,
the average synthesis time amounts to approximately 17.2\,s.  We
consider these times short enough to be of practical use to control
engineers, and thus affirm RQ1.

There are several cases for which the system
may fail to find a controller. For the naive approach, 
the completeness threshold may be too large, thus causing a
timeout.  On the other hand, the abstraction-based approach may
require a very precise abstraction, resulting in too many refinements
and, consequently, a timeout.  Yet another source of incompleteness is the inability
of the {\sc synthesize} phase to use a large enough precision for the
plant model in a given problem.

@@

We have derived a number of benchmarks from the control literature, which we report below. 
As discussed, in this paper, we disregard the presence of output signals, that is the effect of matrices $\textbf{C}, \textbf{D}$, and use
direct state-feedback. 
%
